---
title: "01 · Standard Normal"
subtitle: "The calibration case: exact normality for every n"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("standard_normal"); rows <- population_rows(spec$id)
```

## Research Question

How well do simulation diagnostics behave when the population is already Normal?

## Population / Process and Theory

`X_i ~ N(0, 1)`. Because Normal distributions are closed under averaging, the sample mean is exactly `N(0, 1/n)` for every positive integer `n`, including `n = 1`.

## CLT Assumptions and Prediction

The observations are independent, identically distributed, and have finite variance. I expect all diagnostics to fluctuate around their theoretical targets at every tested `n`; those fluctuations quantify Monte Carlo noise.

## Simulation Design, Search, and Evidence

The standard runner uses `B = 10,000`, the shared coarse grid, and any reviewed refinement values entered in `R/config.R`.

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  plot.new(); text(0.5, 0.5, "Run scripts/run_project.R after entering student_seed.")
} else {
  display_n <- max(rows$n)
  means_path <- file.path(root, "results", "sample_means", paste0(spec$id, "_n_", display_n, ".rds"))
  if (file.exists(means_path)) {
    means <- readRDS(means_path); old_par <- par(mfrow = c(1, 2))
    plot_sampling_histogram(means, display_n, spec$label, 0, 1 / sqrt(display_n))
    plot_sampling_qq(means, display_n, spec$label); par(old_par)
  }
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

The choice `n = 1` is theoretical, not a claim that simulation detected a transition: a Normal population has an exactly Normal sample mean for every positive `n`.
