---
title: "Standard Normal"
author: "Shayaan Tanveer"
---

```{r}
#| include: false
source("R/config.R")
source("R/distributions.R")
source("R/diagnostics.R")
source("R/simulate_means.R")
source("R/plotting.R")
source("R/report_helpers.R")
id <- "standard_normal"
```

## Population and theory

iid N(0,1); the sample mean is exactly N(0,1/n) for every n.

## Expectation based on theory

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Original population shape; this is not the distribution of the sample mean."
plot_population(id)
```

Exact Normal behavior is expected even at n = 1. This expectation was documented during revision; I am not claiming it was a personal prediction made before the original run.

## Evidence for my decision

### Important code, in plain language

The generator calls rnorm(n). Independence and Normality imply that each mean is exactly Normal with mean 0 and SE 1/sqrt(n).

The shared simulator repeats mean(generator(n)) B times and stores one number per sample. Diagnostics are calculated from that vector, not from the original observations. Both seeds save to different directories, and loading a plotted vector checks its mean, SD and Q-Q RMSE against the matching CSV row.

I generate B = 10,000 separate samples at each n using seed 7616. Seed 7617 checks sensitivity. The full tested grid is in the CSV; this table focuses on candidates for our practical-normality review.

```{r}
#| results: asis
review_table(id)
```

```{r}
#| fig-width: 11
#| fig-height: 6
#| fig-cap: "Primary seed 7616. Orange curves use valid theoretical moments. Q-Q steps reflect discreteness; systematic curvature indicates shape differences."
decision <- selected_row(id)
ns <- na.omit(c(decision$below_n, decision$selected_n, decision$above_n))
plot_review_evidence(id, ns)
```

## My selected n and justification

```{r}
#| results: asis
report_selection(id)
```

Here the smallest possible n = 1 is exact by theory, not borderline. Small nonzero simulated diagnostics are sampling noise, not evidence against the theorem.

### Normal calibration

```{r}
calibration <- read.csv("results/normal_benchmark_summary.csv")
knitr::kable(calibration[calibration$n %in% c(1,10,30), ], digits = 3)
```

These empirical 2.5th-97.5th percentile ranges summarize 30 calibration experiments per n for skewness, excess kurtosis and Q-Q RMSE. They contextualize noise; with only 30 experiments they are approximate ranges, not exact tolerance limits.

## Seed sensitivity

```{r}
a <- population_rows(id)
b <- read.csv("results/sensitivity_diagnostics.csv")
b <- b[b$population_id == id, ]
compare <- merge(a[, c("n", "qq_rmse", "skewness")],
                 b[, c("n", "qq_rmse", "skewness")], by = "n",
                 suffixes = c("_7616", "_7617"))
knitr::kable(compare[compare$n %in% sort(unique(c(ns, review_n_values[[id]]))), ], digits = 3)
```

Both seeds validate the implementation, while exact diagnostics change. The selected n = 1 remains exact by theory; no seed choice is needed to justify it.
