---
title: "Six-Sided Die"
author: "Shayaan Tanveer"
---

```{r}
#| include: false
source("R/config.R")
source("R/distributions.R")
source("R/diagnostics.R")
source("R/simulate_means.R")
source("R/plotting.R")
source("R/report_helpers.R")
id <- "six_sided_die"
```

## Population and theory

Independent uniform draws from 1 through 6; mean 3.5, variance 35/12, SE sqrt((35/12)/n).

## Expectation based on theory

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Original population shape; this is not the distribution of the sample mean."
plot_population(id)
```

Symmetry and bounded values suggest relatively fast convergence. This expectation was documented during revision; I am not claiming it was a personal prediction made before the original run.

## Evidence for my decision

### Important code, in plain language

The generator uses sample(1:6, n, replace=TRUE). Replacement makes observations independent uniform die rolls, with mean 3.5 and variance 35/12.

The shared simulator repeats mean(generator(n)) B times and stores one number per sample. Diagnostics are calculated from that vector, not from the original observations. Both seeds save to different directories, and loading a plotted vector checks its mean, SD and Q-Q RMSE against the matching CSV row.

I generate B = 10,000 separate samples at each n using seed 7616. Seed 7617 checks sensitivity. The full tested grid is in the CSV; this table focuses on candidates for our practical-normality review.

```{r}
#| results: asis
review_table(id)
```

```{r}
#| fig-width: 11
#| fig-height: 6
#| fig-cap: "Primary seed 7616. Orange curves use valid theoretical moments. Q-Q steps reflect discreteness; systematic curvature indicates shape differences."
decision <- selected_row(id)
ns <- na.omit(c(decision$below_n, decision$selected_n, decision$above_n))
plot_review_evidence(id, ns)
```

## My selected n and justification

```{r}
#| results: asis
report_selection(id)
```

I use the practical standard approved during review: small remaining bends and steps are acceptable, not exact tail accuracy. Correct mean and SE alone cannot establish a Normal shape. Borderline neighbors mean that the exact minimum is uncertain.

## Seed sensitivity

```{r}
a <- population_rows(id)
b <- read.csv("results/sensitivity_diagnostics.csv")
b <- b[b$population_id == id, ]
compare <- merge(a[, c("n", "qq_rmse", "skewness")],
                 b[, c("n", "qq_rmse", "skewness")], by = "n",
                 suffixes = c("_7616", "_7617"))
knitr::kable(compare[compare$n %in% sort(unique(c(ns, review_n_values[[id]]))), ], digits = 3)
```

The selected value is still a practical candidate under both seeds, although exact diagnostics change. Nearby candidates can be borderline: the scientific interpretation is more stable than an exact boundary integer. I do not choose the seed with the nicest plot.
