---
title: "02 · Six-Sided Die"
subtitle: "A bounded, discrete, symmetric population"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("six_sided_die"); rows <- population_rows(spec$id)
```

## Research Question

How quickly does averaging a discrete uniform die produce a sampling distribution that is comfortable to call approximately normal?

## Population / Process and Theory

`X_i` is uniform on `{1, 2, 3, 4, 5, 6}`. The theoretical mean is `3.5`, variance is `35/12`, and the theoretical SE is `sqrt(35/12) / sqrt(n)`.

## CLT Assumptions and Prediction

The draws are iid with finite variance, so the ordinary CLT applies. Symmetry should reduce skewness, while discreteness may remain visible at small `n`.

## Simulation Design, Search, and Evidence

The runner uses `B = 10,000`, a broad grid, and optional refinements around a reviewed transition. The display uses the largest generated `n` only as a reproducible diagnostic view; it is not a selected N.

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  plot.new(); text(0.5, 0.5, "Run scripts/run_project.R after entering student_seed.")
} else {
  display_n <- max(rows$n); means_path <- file.path(root, "results", "sample_means", paste0(spec$id, "_n_", display_n, ".rds"))
  if (file.exists(means_path)) {
    means <- readRDS(means_path); old_par <- par(mfrow = c(1, 2))
    plot_sampling_histogram(means, display_n, spec$label, spec$theoretical_mean, theoretical_se_for(spec, display_n))
    plot_sampling_qq(means, display_n, spec$label); par(old_par)
  }
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

The die is bounded, symmetric, and finite-variance, so it converges much faster than strongly skewed or heavy-tailed populations. The selected value is the first reviewed value whose graphical and numerical evidence was persuasive together with nearby stability.

Use symmetry, discreteness, benchmark-calibrated diagnostics, theoretical SE agreement, and stability at neighboring values together. Do not infer the answer from one p-value.
