---
title: "03 · Exponential"
subtitle: "A strongly right-skewed finite-variance population"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("exponential"); rows <- population_rows(spec$id)
```

## Research Question

How much averaging is needed to overcome the strong right skew of an Exponential(rate = 1) population?

## Population / Process and Theory

`X_i ~ Exponential(rate = 1)`, with mean `1`, standard deviation `1`, and theoretical SE `1 / sqrt(n)`.

## CLT Assumptions and Prediction

The observations are iid and have finite variance, so the ordinary CLT applies. I expect the right tail and positive skewness to be visible for small `n`, with empirical SE approaching `1 / sqrt(n)` as `n` grows.

## Simulation Design, Search, and Evidence

Use `B = 10,000`. Inspect the broad grid first, then enter a transition center in `R/config.R` to add nearby values.

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  plot.new(); text(0.5, 0.5, "Run scripts/run_project.R after entering student_seed.")
} else {
  display_n <- max(rows$n); means_path <- file.path(root, "results", "sample_means", paste0(spec$id, "_n_", display_n, ".rds"))
  if (file.exists(means_path)) {
    means <- readRDS(means_path); old_par <- par(mfrow = c(1, 2))
    plot_sampling_histogram(means, display_n, spec$label, 1, 1 / sqrt(display_n))
    plot_sampling_qq(means, display_n, spec$label); par(old_par)
  }
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

The selected value is deliberately conservative: `n = 30` still has noticeable positive skewness, while the `n = 1,000` row has near-zero residual shape diagnostics, Q-Q error in the calibrated range, and empirical SE close to `1 / sqrt(n)`.
