---
title: "04 · Binomial Small"
subtitle: "Discrete and skewed: Binomial(size = 5, p = 0.10)"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("binomial_small"); rows <- population_rows(spec$id)
```

## Research Question

Does the CLT behave differently when each observation is a small, highly discrete Binomial count with a low success probability?

## Population / Process and Theory

`Y_i ~ Binomial(size = 5, p = 0.10)`. The Binomial parameter `size = 5` is not the CLT sample size `n`. Here `E[Y_i] = 0.5`, `Var(Y_i) = 0.45`, and the iid theoretical SE is `sqrt(0.45) / sqrt(n)`.

## CLT Assumptions and Prediction

The draws are iid with finite variance, so the ordinary CLT applies. The low mean and discreteness may keep the sampling distribution lumpy and right-skewed at modest `n`.

## Simulation Design and Evidence

The runner uses `B = 10,000`, a broad grid, and reviewed refinements. The display uses the largest generated `n` only to make the report useful after a run.

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  plot.new(); text(0.5, 0.5, "Run scripts/run_project.R after entering student_seed.")
} else {
  display_n <- max(rows$n); means_path <- file.path(root, "results", "sample_means", paste0(spec$id, "_n_", display_n, ".rds"))
  if (file.exists(means_path)) {
    means <- readRDS(means_path); old_par <- par(mfrow = c(1, 2))
    plot_sampling_histogram(means, display_n, spec$label, 0.5, sqrt(0.45) / sqrt(display_n))
    plot_sampling_qq(means, display_n, spec$label); par(old_par)
  }
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

The selected value reflects the rare-event discreteness of `Binomial(size = 5, p = 0.10)`. Here Binomial `size = 5` is a distribution parameter; it is not the CLT sample size `n`. The theoretical SE is `sqrt(0.45 / n)` and agrees with the generated `n = 1,000` result.
