---
title: "Cauchy"
author: "Shayaan Tanveer"
---

```{r}
#| include: false
source("R/config.R")
source("R/distributions.R")
source("R/diagnostics.R")
source("R/simulate_means.R")
source("R/plotting.R")
source("R/report_helpers.R")
id <- "cauchy"
```

## Population and theory

iid standard Cauchy; population mean and variance do not exist. The average remains standard Cauchy for every n.

One theoretical check uses the characteristic function: phi(t) = exp(-abs(t)). Independence gives phi(t/n)^n = exp(-abs(t)), the same characteristic function for the average. Thus no finite n produces convergence to Normal.

## Expectation based on theory

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Original population shape; this is not the distribution of the sample mean."
plot_population(id)
```

Larger n will not produce Normal sample means. This expectation was documented during revision; I am not claiming it was a personal prediction made before the original run.

## Evidence for my decision

### Important code, in plain language

rcauchy(n) generates independent standard Cauchy observations. I calculate finite sample averages but never assign them nonexistent population mean or variance targets.

The shared simulator repeats mean(generator(n)) B times and stores one number per sample. Diagnostics are calculated from that vector, not from the original observations. Both seeds save to different directories, and loading a plotted vector checks its mean, SD and Q-Q RMSE against the matching CSV row.

I generate B = 10,000 separate samples at each n using seed 7616. Seed 7617 checks sensitivity. The full tested grid is in the CSV; this table focuses on candidates for our practical-normality review.

```{r}
#| results: asis
review_table(id)
```

```{r}
#| fig-width: 11
#| fig-height: 6
#| fig-cap: "Primary seed 7616. Orange curves use valid theoretical moments. Q-Q steps reflect discreteness; systematic curvature indicates shape differences."
decision <- selected_row(id)
ns <- na.omit(c(decision$below_n, decision$selected_n, decision$above_n))
plot_review_evidence(id, ns)
```

## Special evidence

```{r}
set.seed(student_seed)
plot_cauchy_running_mean(5000L)
```

## My selected n and justification

```{r}
#| results: asis
report_selection(id)
```

This is a theoretical failure, not an uncertain boundary judgment. Empirical skewness, kurtosis and SD in the table are finite-vector descriptions with no population-moment targets. Standardizing by an outlier-dominated empirical SD can compress the central Q-Q points; it does not rescue Normality.

## Seed sensitivity

```{r}
a <- population_rows(id)
b <- read.csv("results/sensitivity_diagnostics.csv")
b <- b[b$population_id == id, ]
compare <- merge(a[, c("n", "qq_rmse", "skewness")],
                 b[, c("n", "qq_rmse", "skewness")], by = "n",
                 suffixes = c("_7616", "_7617"))
knitr::kable(compare[compare$n %in% sort(unique(c(ns, review_n_values[[id]]))), ], digits = 3)
```

The empirical diagnostics vary substantially between seeds, but the theoretical conclusion does not: no finite n. I do not choose a favorable seed to make a heavy-tailed vector look better.
