---
title: "06 · Extra Credit: Dependent Failure Process"
subtitle: "When covariance changes the variance of a sample mean"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("dependent_failure"); rows <- population_rows(spec$id)
```

## Research Question

How does the sampling distribution of a mean change when each operating time depends on the previous operating time?

## Population / Process and Theory

The instructor model sets the next failure rate to `baseline_rate * exp(beta * (previous_time - reference_time))`, then draws the next operating time conditionally from an Exponential distribution. The supplied values are baseline rate `1`, `beta = 1.25`, and reference time `1`. A long operating period tends to be followed by a shorter one, creating compensating dependence.

## CLT Assumptions and Prediction

Observations are not independent and are not one fixed iid exponential distribution. The variance of a mean includes covariance terms, so the iid `sigma / sqrt(n)` formula is not used as if it were valid.

## Process Diagnostics

```{r}
set.seed(student_seed) # uses the configured student seed for the diagnostic path
failure_path <- simulate_dependent_failure(1000L)
plot_dependent_process(failure_path)
cat("Illustration lag-1 correlation:", cor(failure_path[-length(failure_path)], failure_path[-1L]), "\n")
```

## Simulation Design, Numerical Evidence, and Boundary Search

The mean simulations preserve the dependence inside every generated sample. No independent redraw is substituted. Generated population rows and Q-Q diagnostics are displayed below when available.

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  cat("Run scripts/run_project.R after entering student_seed.\n")
} else {
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

Through `n = 1,000`, the process improves but still has positive skewness and Q-Q error above the Normal calibration. The `N/A` is therefore a tested-range decision, not a claim that no dependent-process CLT could ever be proved. The lag plot and negative lag-1 association also show why an iid standard-error shortcut is unsafe.
