---
title: "Dependent Machine Failure"
author: "Shayaan Tanveer"
---

```{r}
#| include: false
source("R/config.R")
source("R/distributions.R")
source("R/diagnostics.R")
source("R/simulate_means.R")
source("R/plotting.R")
source("R/report_helpers.R")
id <- "dependent_failure"
```

## Population and theory

First rate is 1; later rates are exp(1.25*(previous time - 1)). Each next time is conditionally Exponential. Var(Xbar) includes covariance terms, so an iid SE is inappropriate.

$$
\operatorname{Var}(\bar X)=\frac{1}{n^2}\left[\sum_i\operatorname{Var}(X_i)+2\sum_{i<j}\operatorname{Cov}(X_i,X_j)\right].
$$

Times are nonnegative, so rates are at least exp(-1.25) and conditional second moments are bounded by 2*exp(2.5). Individual finite variance is not enough to prove a dependent CLT. The supplied process starts at rate 1, without a hidden burn-in; its time-specific marginal distributions need not be identical. Compensating dependence may reduce variance relative to an iid shortcut, rather than always increasing it.

## Expectation based on theory

Compensating dependence changes the mean's variance and shape. This expectation was documented during revision; I am not claiming it was a personal prediction made before the original run.

## Evidence for my decision

### Important code, in plain language

The generator first draws at rate 1, then loops forward. Each new rate is computed from the previous operating time before drawing the next time. Restarting this generator for each repetition preserves within-sample dependence and independence between repetitions.

The shared simulator repeats mean(generator(n)) B times and stores one number per sample. Diagnostics are calculated from that vector, not from the original observations. Both seeds save to different directories, and loading a plotted vector checks its mean, SD and Q-Q RMSE against the matching CSV row.

I generate B = 10,000 separate samples at each n using seed 7616. Seed 7617 checks sensitivity. The full tested grid is in the CSV; this table focuses on candidates for our practical-normality review.

```{r}
#| results: asis
review_table(id)
```

```{r}
#| fig-width: 11
#| fig-height: 6
#| fig-cap: "Primary seed 7616. Orange curves use valid theoretical moments. Q-Q steps reflect discreteness; systematic curvature indicates shape differences."
decision <- selected_row(id)
ns <- na.omit(c(decision$below_n, decision$selected_n, decision$above_n))
plot_review_evidence(id, ns)
```

## Special evidence

```{r}
set.seed(student_seed)
path <- simulate_dependent_failure(1000L)
plot_dependent_process(path)
cat("Lag-1 correlation:", cor(head(path, -1), tail(path, -1)))
```

## My selected n and justification

```{r}
#| results: asis
report_selection(id)
```

I use the practical standard approved during review: small remaining bends and steps are acceptable, not exact tail accuracy. Correct mean and SE alone cannot establish a Normal shape. Borderline neighbors mean that the exact minimum is uncertain.

## Seed sensitivity

```{r}
a <- population_rows(id)
b <- read.csv("results/sensitivity_diagnostics.csv")
b <- b[b$population_id == id, ]
compare <- merge(a[, c("n", "qq_rmse", "skewness")],
                 b[, c("n", "qq_rmse", "skewness")], by = "n",
                 suffixes = c("_7616", "_7617"))
knitr::kable(compare[compare$n %in% sort(unique(c(ns, review_n_values[[id]]))), ], digits = 3)
```

The selected value is still a practical candidate under both seeds, although exact diagnostics change. Nearby candidates can be borderline: the scientific interpretation is more stable than an exact boundary integer. I do not choose the seed with the nicest plot.
