---
title: "Non-Identical Basketball Shots"
author: "Shayaan Tanveer"
---

```{r}
#| include: false
source("R/config.R")
source("R/distributions.R")
source("R/diagnostics.R")
source("R/simulate_means.R")
source("R/plotting.R")
source("R/report_helpers.R")
id <- "nonidentical_bernoulli"
```

## Population and theory

Independent Bernoulli shots with probabilities evenly spanning 0.90 to 0.50. Mean is 0.70 for n >= 2, and 0.90 at n = 1. Var(Xbar) = sum p_i(1-p_i)/n^2.

## Expectation based on theory

Bounded independent observations can support a generalized CLT despite changing probabilities. This expectation was documented during revision; I am not claiming it was a personal prediction made before the original run.

## Evidence for my decision

### Important code, in plain language

seq(0.90, 0.50, length.out=n) creates the probabilities. rbinom(n, size=1, prob=probabilities) draws independent shots with those different probabilities. The SE helper sums their individual variances and divides by n squared before taking a square root.

The shared simulator repeats mean(generator(n)) B times and stores one number per sample. Diagnostics are calculated from that vector, not from the original observations. Both seeds save to different directories, and loading a plotted vector checks its mean, SD and Q-Q RMSE against the matching CSV row.

I generate B = 10,000 separate samples at each n using seed 7616. Seed 7617 checks sensitivity. The full tested grid is in the CSV; this table focuses on candidates for our practical-normality review.

```{r}
#| results: asis
review_table(id)
```

```{r}
#| fig-width: 11
#| fig-height: 6
#| fig-cap: "Primary seed 7616. Orange curves use valid theoretical moments. Q-Q steps reflect discreteness; systematic curvature indicates shape differences."
decision <- selected_row(id)
ns <- na.omit(c(decision$below_n, decision$selected_n, decision$above_n))
plot_review_evidence(id, ns)
```

## Special evidence

```{r}
plot(nonidentical_probabilities(100L), type = "l", xlab = "Shot", ylab = "Probability")
```

Each n gets a new evenly spaced probability vector; it is not a prefix of one fixed 100-shot sequence. For n >= 2 the mean probability is 0.70, but pretending every shot has p = 0.70 gives the wrong variance.

```{r}
n <- selected_row(id)$selected_n
p <- nonidentical_probabilities(n)
knitr::kable(data.frame(n = n,
  expected_mean = mean(p),
  correct_SE = sqrt(sum(p*(1-p))/n^2),
  iid_shortcut_SE = sqrt(mean(p)*(1-mean(p))/n)), digits = 4)
```

The difference is mathematical: average[p_i(1-p_i)] equals p-bar(1-p-bar) minus average[(p_i-p-bar)^2]. Therefore the iid shortcut overstates the variance here.

## My selected n and justification

```{r}
#| results: asis
report_selection(id)
```

I use the practical standard approved during review: small remaining bends and steps are acceptable, not exact tail accuracy. Correct mean and SE alone cannot establish a Normal shape. Borderline neighbors mean that the exact minimum is uncertain.

## Seed sensitivity

```{r}
a <- population_rows(id)
b <- read.csv("results/sensitivity_diagnostics.csv")
b <- b[b$population_id == id, ]
compare <- merge(a[, c("n", "qq_rmse", "skewness")],
                 b[, c("n", "qq_rmse", "skewness")], by = "n",
                 suffixes = c("_7616", "_7617"))
knitr::kable(compare[compare$n %in% sort(unique(c(ns, review_n_values[[id]]))), ], digits = 3)
```

The selected value is still a practical candidate under both seeds, although exact diagnostics change. Nearby candidates can be borderline: the scientific interpretation is more stable than an exact boundary integer. I do not choose the seed with the nicest plot.
