---
title: "07 · Extra Credit: Non-Identical Bernoulli Shots"
subtitle: "Independent shots with probabilities that change from 0.90 to 0.50"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("nonidentical_bernoulli"); rows <- population_rows(spec$id)
```

## Research Question

Does a normal approximation emerge similarly when observations remain independent but are not identically distributed?

## Population / Process and Theory

For each candidate `n`, probabilities `p_i` are evenly spaced from 0.90 to 0.50, and `X_i ~ Bernoulli(p_i)` independently. The theoretical mean of the average is `(1/n) * sum(p_i)`. The variance is `sum(p_i * (1 - p_i)) / n^2`, not an iid Binomial calculation.

## CLT Assumptions and Prediction

Independence and finite variance hold, while identical distribution does not. Since the probabilities vary smoothly and stay away from 0 and 1, a normal approximation may be reasonable, but the relevant mean and variance change with candidate `n`.

## Probability Curve

```{r}
probabilities <- nonidentical_probabilities(20L)
plot(seq_along(probabilities), probabilities, type = "b", pch = 19,
     col = "#C44E52", ylim = c(0, 1), xlab = "Shot number",
     ylab = "P(make)", main = "Basketball success probability by shot")
```

## Simulation Design and Evidence

Each simulated sample calls `rbinom(n, size = 1, prob = p)` with its own probability vector. The displayed diagnostics use the correct `p_i`-specific theoretical SE.

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  plot.new(); text(0.5, 0.5, "Run scripts/run_project.R after entering student_seed.")
} else {
  display_n <- max(rows$n); means_path <- file.path(root, "results", "sample_means", paste0(spec$id, "_n_", display_n, ".rds"))
  if (file.exists(means_path)) {
    means <- readRDS(means_path); old_par <- par(mfrow = c(1, 2)); theory <- nonidentical_bernoulli_theory(display_n)
    plot_sampling_histogram(means, display_n, spec$label, theory$mean, theory$sd)
    plot_sampling_qq(means, display_n, spec$label); par(old_par)
  }
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

The selected value uses `Var(Xbar) = sum(p_i * (1 - p_i)) / n^2`; replacing the changing probabilities by one iid Bernoulli variance would answer a different problem. The `n = 1,000` row is numerically stable against the alternate seed, although convergence is slower than for the bounded iid die.
