---
title: "08 · Extra Credit: Synthetic Age-at-Death"
subtitle: "An irregular discrete mixture with early and late peaks"
---

```{r}
root <- if (file.exists("R/config.R")) "." else ".."
source(file.path(root, "R/config.R")); source(file.path(root, "R/distributions.R"))
source(file.path(root, "R/diagnostics.R")); source(file.path(root, "R/simulate_means.R"))
source(file.path(root, "R/plotting.R")); source(file.path(root, "R/report_helpers.R"))
spec <- population_spec("age_at_death"); rows <- population_rows(spec$id)
```

## Research Question

How does an irregular, bimodal-looking synthetic population behave after averaging?

## Population / Process and Theory

The instructor construction uses ages 0 through 100 with weights `0.045 * exp(-x/5) + 0.030 * exp(-0.5 * ((x - 65) / 10)^2)`, normalized to probabilities. The supplied `rdeath()` logic samples from those probabilities. The resulting distribution is synthetic, not historical mortality data; its mean and SD are computed from the normalized table.

## CLT Assumptions and Prediction

The draws are iid with finite variance. The unusual shape may make small-n sampling distributions depart from normality even though ordinary CLT conditions hold. I expect sample means to smooth the irregular population as `n` increases.

## Original Population and Simulation Evidence

```{r}
distribution <- age_at_death_distribution()
plot(distribution$age, distribution$percent, type = "l", lwd = 3,
     col = "#C44E52", xlab = "Age at death", ylab = "Probability (%)",
     main = "Synthetic age-at-death population")
```

```{r}
if (is.null(rows) || nrow(rows) == 0L) {
  cat("Run scripts/run_project.R after entering student_seed.\n")
} else {
  print(rows)
}
```

## Final Selected N and Interpretation

```{r, results='asis'}
report_selection(spec$id)
```

The selected value reflects the synthetic two-component discrete mixture supplied by the instructor. It is not historical mortality data; the irregular source shape is exactly what makes the fast stabilization of the sample mean worth checking.
